The Dynamics of Anthropomorphic Effects in Human-Robot Interaction

### 2014-03-24 paper.tex
modifications Ranger domino study, structure anthropomorphism section

\usepackage{graphicx}
\graphicspath{{figs/}}
\usepackage{tikz}
\usetikzlibrary{shapes}
\usetikzlibrary{svg.path}
%
\usepackage{amsmath} 
\usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
\usepackage{color, soul}
\usepackage{url}
\usepackage{multirow}
\usepackage{array}
\usepackage{fixltx2e}
\usepackage{textcomp}

@@

\usepackage{xifthen}% provides \isempty test

% use either \Ant or \Ant[R,H,C,...]
\newcommand{\Ant}[1][]{%
      \ifthenelse{\isempty{#1}}%
        {$\mathcal{A}$}
        {$\mathcal{A}(#1)$}
}

% for use inside a math environment
\newcommand{\ant}[1][]{%
      \ifthenelse{\isempty{#1}}%
        {\mathcal{A}}
        {\mathcal{A}(#1)}
}
\newcommand{\ICA}{{$\mathcal{A}_0$~}}
\newcommand{\SLA}{{$\mathcal{A}_\infty$~}}
\newcommand{\sla}{{\mathcal{A}_\infty}}
\newcommand{\AntMax}{{$\mathcal{A}_{max}$~}}

@@

While anthropomorphism in robotics is a commonly discussed trait of
HRI, it paradoxically lacks formal grounds. This article
attempts first at providing a comprehensive synthesis of the social
phenomenon of \textit{perceiving human-like characteristics in non-human
agents and attributing those characteristics to robots}. We draw on
social sciences and psychology, as well as on a critical survey of
existing literature in the HRI community, to ground the concept of
\textit{anthropomorphism}. Supported by extensive literature review
and long-term field studies, we suggest to go beyond the traditional
perception of anthropomorphism as a static feature of a system: we propose to
understand anthropomorphism as a dynamic, non-monotonic and context-dependent
process, which evolves over time and accounts for special events such as the
so-called \textit{novelty effect}. To this end, we introduce a model of
anthropomorphism that analyzes the phenomenon along three interaction phases,
and we discuss cognitive interpretations that follow. \hl{(Julia: probably the abstract needs to be refined once we finished writing the paper)}

@@

A phenomenon that is often studied related to the human-like design of robots is
\textbf{anthropomorphism} (noted \Ant in this paper). Besides a feeling of affinity, and empathy,
anthropomorphism can be understood as being one of the resulting (and  more or
less desired) effects of the anthropomorphic design of artificial agents (up to
the first positive peak). The essence of anthropomorphism is the human user's
perception of human-likeness in the non-human agent. Logically, anthropomorphism
is said to be enhanced with increased human-like design of the agent. There are
various other reasons and explanations for why people tend to anthropomorphize
non-human agents. In general, as anthropomorphism emerges in a (real or
imagined) interaction \citep{persson_anthropomorphism_2000} between at least two
players (the human who anthropomorphizes and the non-human agent which is
anthropomorphized) there are also two main perspectives on the phenomenon: an
artifact-centered (design) and a human-centered (cognitive, psychological) one.
The Uncanny Valley hypothesis can be categorized as being part of first,
artifact-centered approach, to understand anthropomorphism.\\

We observed anthropomorphism in several short- and long-term human-robot
interaction studies that we carried out with embodied robots, mostly in
real-world settings. There is not much research that investigates
anthropomorphism based on real interaction between a human and a robot, and
above all in natural settings and over extended periods of time.

Based on these observations and an extensive literature review, we would like to
propose a model that accounts for variances and dynamics in anthropomorphism.
This model of the \textit{``Dynamics of Anthropomorphism''} considers several
impact factors, including the human user, the anthropomorphic design of the
non-human agent, the context of use, and the interaction experience (time). With
this initial model, we would like to extend the current understanding of
anthropomorphism in robotics and human-robot interaction. This might in turn
lead to a discussion about the short- and long-term benefits and other effects
of the human-like (anthropomorphic) design of robots and other artificial
agents.\\

@@

To support the reflection, it may be useful to consider a few experimental
results. Figure~\ref{roomba-perception} gives a first perspective on how naive
users perceive robots over a long period of time. Here, \emph{Roomba} robots
were given to nine household for a 6 months period. The aim was to explore how
people use and perceive the robot over time, and how they generally live with
the Roomba in their home. At various time points, users were asked to rated
their perception of different features of the robot on a 7-Likert scale. As
Figure~\ref{roomba-perception} shows, some features remained stable over time
(\emph{ease of use} and \emph{fun}), and other decreased: the perceived
\emph{usefulness}, \emph{intelligence} and \emph{impact} on their household.
\emph{intelligence} and \emph{impact} reflect the cognitive engagement of the
user towards the machine: we can assume that one do not care to engage in a
(cognitive) interaction with an artifact perceived as non-intelligent and with
no impact on his/her life. Even if one can argue that the Roomba is not designed
to foster interaction, the question is raised: how to sustain engagement between
a robot and a human over a long period of time? And, as a pre-requisite, do we
actually understand the psychological phenomenon, and can we effectively account
for the dynamics of these interactions?

Those questions are certainly not new, but paradoxically, while anthropomorphism
in robotics is a commonly discussed and studied trait of human-robot
interaction, it appears it lacks a formal ground and unified understanding to
take into account the wide range of phenomena that it encompasses.

@@

\fxnote{state here our main research hypothesis}
\fxnote{state here our main contributions}

So far, the HRI community has not much investigated how anthropomorphism in
human-robot interactions evolves over time (during the process of
\emph{adopting} a robot, for instance). Anthropomorphism is traditionally
perceived in robotics as a static feature that once observed during a
short-term interaction reflects a sustaining social effect.  Based on an
literature review previously published~\cite{fink_anthropomorphism_2012}, a
long-term field study in a natural environment~\cite{fink_living_2013}, as well
as two on-going child-robot experiments, we believe that anthropomorphic
effects do evolve over time in non-monotonic ways, and play a central role in
sustaining engagement in human-robot interaction.

We propose to formalize these dynamic effects into a model that we call the
\emph{dynamics of anthropomorphism}. We hope that it may support further
discussions and reflections on how anthropomorphism impacts human-robot
interaction on the long run, and also foster research on the affective bonds
induced by anthropomorphic projections on robots.

As such, we propose to represent how the level of anthropomorphic effects (\ie
observable manifestations of anthropomorphism) evolves over a long-term
human-robot interaction (Figure~\ref{fig:dynamics}). By long-term interaction,
we mean direct (non-mediated), repeated interaction with the same robot, over
an extended period of time (typically longer than a week).

In this article, we would like to share some critical but constructive
considerations on anthropomorphism in robotics and HRI. Thereby, we contribute
to a better understanding of this complex phenomenon related to robots.

The remainder of this article is organized as follows.  First, in Section
\ref{sec:overview-studies}, we give a brief overview on the various studies that
we conducted to investigate anthropomorphism, namely (1) the Roomba study, (2)
the Forum study, (3) the Ranger domino study. As an outcome of these studies, we
propose two main research hypotheses related to anthropomorphism.  Next, in
Section \ref{sec:anthropomorphism} we investigate these hypotheses by
synthesizing literature on anthropomorphism. This should also serve to establish
a common ground. We first explore an artifact-centered perspective and briefly
outline how the human-like design of agents can encourage anthropomorphism.
Then, we explore a human-centered perspective by integrating theories from
cognitive science, developmental and social psychology that serve to explain
anthropomorphism. At the end of this section we bring both perspectives together
and interpret that anthropomorphism is a multi-factor phenomenon.  Based on the
two previous sections, we propose an initial model of the \textit{``Dynamics of
Anthropomorphism''} in Section \ref{sec:our-ideas}. This model describes the
variances and dynamics of people's tendency to anthropomorphize an artificial
agent, such as a robot.  Next, in Section \ref{sec:cognition-neuroscience} we
present a cognitive and neuroscience perspective on anthropomorphism to
consolidate our considerations and the proposed model.  This will be followed by
a discussion about the limitations of our approach (Section
\ref{sec:discussion}). Further, we debate existing and alternative ways to
measure anthropomorphism in human-robot interactions.  Finally, Section
\ref{sec:conclusion} concludes this article and offers starting points for
further research on anthropomorphism and the effect of the human-like design
approach in robotics.

@@

\section{Studying anthropomorphism}
\label{sec:overview-studies}

@@

We openly explored long-term interactions between human users and a domestic
service robot in a 6-months lasting ethnographic study. We gave a commercially
available vacuum cleaning robot (iRobot's Roomba, see Figure
\ref{fig:aibo-roomba-ipad} middle) to nine households in the French speaking
part of Switzerland. The fairly diverse sample of households consisted of 2
single homes, 1 couple, and 6 families with children in their home. In total,
this summed up to 30 participants: 15 adults (26-71 years old, mean age 43.6
years), and 15 children (6 months to 18 years old). During the 6 months, each
household was visited 5 times to conduct interviews and make field observations
in the households. Moreover, during several weeks, households were asked to fill
in a diary to report the usage of the robot and other cleaning activities.
\footnote{This study is published in \cite{fink_living_2013}.} Hereafter, we
refer to this study as the "Roomba study".

@@

From the Roomba study, we can report on several findings related to
anthropomorphism and the anthropomorphic design of robots.
%The gathered data for these aspects include qualitative observations,
%semi-structured interviews, and also quantitative assessments of various
%aspects of human-robot interaction.
The Roomba study is one of very few long-term human-robot interaction studies in
the domestic environment.

In the following, we present findings in respect of two aspects: First, we
qualitatively describe what kinds of anthropomorphic interactions between
participants and the robot were observed and outline how these interactions
evolved over time.  Second, we present people's perception of human-likeness and
social capabilities of the robot and how this perception changed over
time.\footnote{\hl{note that these two different kinds of data
(behavior/interaction and perception) already show that we try to measure
anthropomorphism from these two angles!}}

@@

It has been described in several previous studies of the Roomba that users tend
to engage in a variety of non-cleaning activities associated with the robot,
including, but not limited to, naming, ascribing personality and gender
\citep{forlizzi_service_2006,forlizzi_how_2007,krumm_my_2007,sung_domestic_2010}.
\cite{sung_housewives_2008} carried out a survey among 379 Roomba users to
investigate how common these activities are. The authors concluded that overall,
respondents had relatively low rates of treating Roomba as a living entity.

In our long-term study of the Roomba in homes, we found that it was mostly
during the initial phase with the robot that some (rather few) people tended to
interact in an anthropomorphic or human-social way with their vacuum cleaning
robot.  Overall, our data suggest that the anthropomorphic effect and the social
impact of a functional robot on a household is rather marginal.
%There were just few instances of people anthropomorphizing the robot and
%interacting with it in a human-social way. More importantly, overall, these
%anthropomorphic interactions with the robot did not last but seemed to be
%somehow related to the novelty of the robot.

We categorized several interactions and actions with the robot as
anthropomorphic or human-social (in a broader sense). 
%Examples of what we categorized as somehow social and anthropomorphic
%(human-like) interactions with the robot include \textit{watching} the robot
%for fun, giving a \textit{demo} of the robot to guests / visitors, taking a
%\textit{photo/video} of the robot, \textit{personalizing} the robot, giving a
%\textit{name} to the robot, \textit{talking} to the robot using direct speech,
%\textit{playing} with the robot, and making \textit{adjustments} to the home
%due to the robot. 
Figure \ref{fig:roomba-activities}) shows how many of the 9 households engaged
in these activities during the 6 months when the study was conducted.

\begin{figure}
    \centering
    \includegraphics[width=0.5\columnwidth]{roomba-activities.png}
    \caption{Activities with the robot over time, n = 9 households. This bar
        chart shows how many of the 9 households carried out different kinds of
        anthropomorphic / social activities with and related to the robot at 3
        time points during 6 months. Activities include \texttt{watching} the
        robot for fun, giving a \texttt{demo} of the robot to guests / visitors,
        \texttt{personalizing} the robot, giving a \texttt{name} to the robot,
        \texttt{talking} to the robot, \texttt{playing} with the robot.
        Participants were asked to report on these activities at 3 time points
        after the robot was deployed in their home: after 2 weeks, after 2 months,
        and after 6 months.}

@@

We also studied people's perception of the robot, and how it evolved over time
(see Figure \ref{fig:roomba-perception}). The main idea was to find out which of
the perceived attributes could be related to how far the robot was integrated in
the household as a cleaning tool. Concerning anthropomorphism, we included
\textit{perceived intelligence} and experienced \textit{fun} when using the
robot as indicators for the perception of human-likeness in the robot.

\begin{figure}
    \centering

    \begin{tikzpicture}[scale=0.5, y = 2cm]
    %axis
    \draw (0,0) -- coordinate (x axis mid) (20,0);
    \draw (0,0) -- coordinate (y axis mid) (0,7);
    %ticks
    \foreach \x in {0,4,...,20}
        \draw (\x,1pt) -- (\x,-3pt);
    \foreach \y in {1,...,7} 
        \draw[black!20] (20,\y) -- (-3pt,\y);
    \foreach \y in {1,...,7} 
        \draw (-3pt, \y) node[anchor=east] {\y}; 
    %labels
    \node[below=0.8cm] at (x axis mid) {People's perception ($\sigma=15$)};
    \node at (2, -0.4) {\tiny expectations};
    \node[align=center] at (6, -0.4) {\tiny after\\ \tiny first usage};
    \node[align=center] at (10, -0.4) {\tiny after\\ \tiny 2 weeks};
    \node[align=center] at (14, -0.4) {\tiny after\\ \tiny 2 months};
    \node[align=center] at (18, -0.4) {\tiny after\\ \tiny 6 months};

    \draw[dashed, black!50] (6,0) -- (6,7) node[sloped, below, pos=0.25] {\tiny novelty effect};

    \draw[ultra thick, red!50] (2,5.6) -- (6,6.1) -- (10,6.3) -- (14,6.3) -- (18,6.1) node at (15,6.5) {\tiny \it ease of use};
    \draw[ultra thick, purple] (2,4.9) -- (6,5.2) -- (10,5.1) -- (14,4.5) -- (18,4.2) node at (3,4.7) {\tiny \it usefulness};
    \draw[ultra thick, blue!50] (2,4) -- (6,4.1) -- (10,4.6) -- (14,4.1) -- (18,4.3) node at (16,4) {\tiny \it fun};
    \draw[ultra thick, green!50] (2,3.9) -- (6,4.1) -- (10,4.2) -- (14,3.8) -- (18,3.3) node at (19,3.5) {\tiny \it intelligence};
    \draw[ultra thick, orange] (2,3.5) -- (6,4.2) -- (10,4) -- (14,3.2) -- (18,2.8) node at (14,2.7) {\tiny \it impact};

    \draw[ultra thick, dashed] (2,5.2) -- (6,5.2) -- (10,5) -- (14,4.9) -- (18,4.5) node at (14,5.3) {\tiny \it overall};

    \end{tikzpicture}

@@

\begin{figure}
    \centering
    \includegraphics[width=0.7\columnwidth]{aibo-roomba-ipad_2.jpg}
    \caption{The Sony AIBO robotic dog (left), the iRobot Roomba vacuum-cleaning robot (middle), and the Apple Inc. iPad tablet computer (right) were compared in terms of how far users of these devices use anthropomorphic expressions in discussions about their technology / robot in online discussion forums.}
    \label{fig:aibo-roomba-ipad}
\end{figure}

@@

    	\item People engage in a more social and human-like way with a robotic pet compared to a functional robot and a tablet computer, therefore there will be a higher amount of anthropomorphic language in the AIBO and Roomba forum, compared to the iPad forum.

@@

@@

\begin{itemize}
    \item \textit{Life-likeness}, uch as being alive or having (parts of) a
        body, \eg \textit{``she can be reborn since you copied her memory
        stick''};

    \item \textit{Emotional states, empathy} or the technology having a
        \textit{feeling}, \eg \textit{``oh no, poor AIBO''};

    \item \textit{Gender, personality} or the technology having an
        \textit{intention}, \eg Roomba \textit{``seems to like to hang around
        under the sofa''};

    \item When the author gave the technology a \textit{name}, \eg
        \textit{``whenever I show Java his pink ball''};

    \item Being \textit{socially integrated}, such as being a family member, \eg
        \textit{``I'm considering adding a Roomba to the family''};

    \item \textit{Metaphorical ways}, \eg Roomba \textit{``sings its victory
        song when it finishes and docks''}.

\end{itemize}

In the next sections, we present findings in respect of our two hypotheses:
anthropomorphic language related to the device, and anthropomorphic language
related to the topic of the forum post.

Overall, our analysis revealed that in all three forums, discussions about
\textit{technical} aspects of the respective device were most frequent (67.7~\%
of all 1208 categorized segments). However there were huge differences between
the three devices: \textit{technical} aspects made up 79.9~\% of the discussion
in the Roomba forum, compared to 69.9~\% in the AIBO forum, and 55.2~\% in the
iPad forum. Another fairly frequently discussed topic concerned the
\textit{usage} of the respective device, thus people describing how and for what
they use their technology. This topic made up 21.8~\% of all posts, and was most
frequent in the iPad forum (35.7~\%), followed by the Roomba forum (15.9~\%),
and the AIBO forum (10.7~\%). Last but not least, we defined \textit{relation}
as the third topic of interest. Discussions in which people described their
relation to the respective device made up 10.5~\% of all posts, and was most
frequent in the AIBO forum (19.4~\%), followed by the iPad forum (9.1~\%), and
the Roomba forum (4.2~\%). \hl{(if it is getting too much we can leave out this
paragraph, it's not crucial, and doesn't say anything about anthropomorphism
directly)}

@@

13 groups of each two children (n = 26) who were all between 4-5 years old and
knew each other, were invited to play for around 20 minutes with the Ranger
robot toy box \citep{fink2014which} and assemble a domino together. The robot
was operated by one of the experimenters using the wizard-of-oz technique.
Ranger was supposed to transport domino tiles between the two children and it
would do its job correctly for the first 7 domino tiles. Then, however, the
robot started misbehaving by either \textit{disobeying} to the child

>>>>

\usepackage{graphicx}
\graphicspath{{figs/}}
\usepackage{tikz}
\usetikzlibrary{shapes}
%
\usepackage{amsmath} 
\usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
\usepackage{color, soul}
\usepackage{url}
\usepackage{multirow}
\usepackage{array}
\usepackage{fixltx2e}
\usepackage{textcomp}

@@

@@

A phenomenon that is often studied related to the human-like design of 
robots is anthropomorphism. While people's tendency to perceive human-like 
characteristics in robots is a commonly discussed trait of 
human-robot interaction (HRI), the phenomenon paradoxically lacks formal 
grounds. Based on findings from several empirical studies and a 
comprehensive synthesis of literature, this article attempts at extending 
the current understanding of anthropomorphism. In accordance with theories 
from social psychology, cognitive science, and neuroscience, we suggest to 
formalize qualitative differences and dynamics over time in how 
humans anthropomorphize non-human agents. Based on several key factors, 
namely the 
design of the non-human agent, the individual characteristics of the human 
user, and the context of use, we found that anthropomorphism changes in a 
non-linear way with growing experience of the user with the system. To 
account for these variances, we propose an initial model that we call 
\textit{``Dynamics of Anthropomorphism''}. Focusing on human-robot 
interactions, we present a cognitive interpretation that considers 
the described dynamics. Finally, we would like to discuss how others can 
apply our model to measure and describe better the complexity in human's 
tendency to perceive human-like characteristics in non-human agents.

@@

A phenomenon that is often studied related to the human-like design of robots is
\textbf{anthropomorphism}. Besides a feeling of affinity, and empathy,
anthropomorphism can be understood as being one of the resulting (and  more or
less desired) effects of the anthropomorphic design of artificial agents (up to
the first positive peak). The essence of anthropomorphism is the human user's
perception of human-likeness in the non-human agent. Logically, anthropomorphism
is said to be enhanced with increased human-like design of the agent. There are
various other reasons and explanations for why people tend to anthropomorphize
non-human agents. In general, as anthropomorphism emerges in a (real or
imagined) interaction \citep{persson_anthropomorphism_2000} between at least two
players (the human who anthropomorphizes and the non-human agent which is
anthropomorphized) there are also two main perspectives on the phenomenon: an
artifact-centered (design) and a human-centered (cognitive, psychological) one.
The Uncanny Valley hypothesis can be categorized as being part of first,
artifact-centered approach, to understand anthropomorphism.

@@

To support the reflection, it may be useful to consider a few experimental
results. We observed anthropomorphism in several short- and long-term human-robot
interaction studies that we carried out with embodied robots, mostly in
real-world settings. There is not much research that investigates
anthropomorphism based on real interaction between a human and a robot, and
above all in natural settings and over extended periods of time.

Figure~\ref{roomba-perception} gives a first perspective on how naive
users perceive robots over a long period of time. Here, \emph{Roomba} vacuum cleaning robots
were given to nine households for a 6-months period. The aim was to explore how
people use and perceive the robot over time, and how they generally live with
the Roomba in their home. At various time points, users were asked to rate
their perception of different features of the robot on a 7-point rating scale. As
Figure~\ref{roomba-perception} shows, some features remained stable over time
(\emph{ease of use} and \emph{fun}), and others decreased: the perceived
\emph{usefulness}, \emph{intelligence} and \emph{impact} on their household.
\emph{intelligence} and \emph{impact} reflect the cognitive engagement of the
user towards the machine: we can assume that one does not care to engage in a
(cognitive) interaction with an artifact perceived as non-intelligent and with
no impact on his/her life. Even if one can argue that the Roomba is not designed
to foster interaction, the question is raised: how to sustain engagement between
a robot and a human over a long period of time? And, as a pre-requisite, do we
actually understand the psychological phenomenon, and can we effectively account
for the dynamics of these interactions?

Those questions are certainly not new, but paradoxically, while anthropomorphism
in robotics is a commonly discussed and studied trait of human-robot
interaction, it appears that it lacks a formal ground and unified understanding to
take into account the wide range of phenomena that it encompasses.

@@

So far, the HRI community has not much investigated how anthropomorphism in
human-robot interactions evolves over time. In robotics, anthropomorphism is traditionally
perceived as a static feature triggered by the human-like design of a robot that once observed (during a
short-term interaction) reflects a sustaining social effect. Based on a previously published
literature review~\citep{fink_anthropomorphism_2012}, a
long-term field study in a natural environment~\citep{fink_living_2013}, a 
content analysis of online discussion 
forums~\citep{fink_anthropomorphic_2012}, and an on-going child-robot 
interaction study, we believe that anthropomorphic
effects are based on several factors, and that they do evolve over time in 
non-monotonic ways. Moreover, we assume that these variances and dynamics 
play a central role in sustaining engagement in human-robot interaction.

This article is organized around two main hypothesis that we formulated based on 
our empirical studies of HRI, and by integrating previous work from several domains.

\begin{itemize}
    \item \textbf{Hypothesis 1~} A person's tendency to anthropomorphize a robot
is based on several factors, including (but not limited to) (1) the 
design of the robot, (2) the personal characteristics of the human user, 
and (3) the context and situation of use. For instance, a social 
context and purpose of the robot increase anthropomorphism. 
\textit{(anthropomorphism as a multi-factor phenomenon)}
	\item \textbf{Hypothesis 2~} A person's tendency to anthropomorphize a robot 
evolves in a non-linear way over time and with growing experience that the 
person has with the robot. For instance, a more experienced user will perceive 
the robot as more familiar and predictable, and in turn anthropomorphizes the 
robot less. \textit{(anthropomorphism as a dynamic phenomenon)}
\end{itemize}

We propose to formalize these variances and dynamic effects into a model that we call the
\emph{Dynamics of Anthropomorphism}. We hope that it may support further
discussions and reflections on how anthropomorphism impacts human-robot
interaction on the long run, and also foster research on the affective bonds
induced by anthropomorphic projections on robots.
Through our model, we propose to represent how the level of anthropomorphic effects (\ie
observable manifestations of anthropomorphism) evolves over a long-term
human-robot interaction (Figure~\ref{fig:dynamics}). By long-term interaction,
we mean direct (non-mediated), repeated interaction with the same robot, over
an extended period of time (typically longer than a week).

%In this article, we would further like to share some critical but constructive
%considerations on anthropomorphism in robotics and HRI. Thereby, we contribute
%to a better understanding of this complex phenomenon related to robots.

The remainder of this article is organized as follows.  First, in Section
\ref{sec:overview-studies}, we give a brief overview on the various studies that
we conducted to investigate anthropomorphism, namely (1) the Roomba study, (2)
the Forum study, (3) the Ranger domino study. Next, in Section 
\ref{sec:anthropomorphism} we aim to establish a common ground in the understanding 
of anthropomorphism. We first explore an artifact-centered perspective and briefly
outline how the human-like design of agents can encourage anthropomorphism.
Then, we explore a human-centered perspective by integrating theories from
cognitive science, developmental and social psychology that serve to explain
anthropomorphism. At the end of this section we bring both perspectives together
and interpret that anthropomorphism is a multi-factor phenomenon.  Based on the
two previous sections, we propose an initial model of the \textit{``Dynamics of
Anthropomorphism''} in Section \ref{sec:our-ideas}. This model describes the
variances and dynamics of people's tendency to anthropomorphize an artificial
agent, such as a robot.  Next, in Section \ref{sec:cognition-neuroscience} we
present a cognitive and neuroscience perspective on anthropomorphism to
consolidate our considerations and the proposed model.  This will be followed by
a discussion about the limitations of our approach (Section
\ref{sec:discussion}). Further, we debate existing and alternative ways to
measure anthropomorphism in human-robot interactions.  Finally, Section
\ref{sec:conclusion} concludes this article and offers starting points for
further research on anthropomorphism and the effect of the human-like design
approach in robotics.

@@

\section{Studying anthropomorphism in HRI}
\label{sec:overview-studies}

@@

We openly explored long-term interactions between human users and a domestic
service robot in a 6-months lasting ethnographic study~\citep{fink_living_2013}. 
We gave a commercially
available vacuum cleaning robot (iRobot's Roomba, see Figure
\ref{fig:aibo-roomba-ipad} middle) to nine households. 
%in the French speaking
%part of Switzerland. 
%The fairly diverse sample of households consisted of 2
%single homes, 1 couple, and 6 families with children in their home. In total,
%this summed up to 30 participants: 15 adults (26-71 years old, mean age 43.6
%years), and 15 children (6 months to 18 years old). During the 6 months, each
%household was visited 5 times to conduct interviews and make field observations
%in the households. Moreover, during several weeks, households were asked to fill
%in a diary to report the usage of the robot and other cleaning activities.
Hereafter, we refer to this study as the ``Roomba study''. It is one of very few 
long-term human-robot interaction studies in the domestic environment. 
In the following, we present findings in respect of two aspects: First, we
qualitatively describe what kinds of anthropomorphic interactions between
participants and the robot were observed and outline how these interactions
evolved over time.  Second, we present people's perception of human-likeness and
social capabilities of the robot and how this perception changed over
time.\footnote{\hl{note that these two different kinds of data
(behavior/interaction and perception) already show that we try to measure
anthropomorphism from these two angles!}}

@@

%The gathered data for these aspects include qualitative observations,
%semi-structured interviews, and also quantitative assessments of various
%aspects of human-robot interaction.

@@

It has been described in several previous studies of the Roomba that users tend
to engage in a variety of non-cleaning activities associated with the robot,
including, but not limited to, naming, ascribing personality and gender
\citep{forlizzi_service_2006,forlizzi_how_2007,krumm_my_2007,sung_domestic_2010}.
\cite{sung_housewives_2008} carried out a survey among 379 Roomba users to
investigate how common these activities are. The authors concluded that overall,
respondents had relatively low rates of treating Roomba as a living entity.  
Similarly, in our long-term study of the Roomba in homes, we found that only few 
people tended to interact in an anthropomorphic or human-social way with their 
vacuum cleaning robot and that this effect did not sustain. Overall, our data 
suggest that the anthropomorphic effect and the social
impact of a functional robot on a household are rather marginal.
%There were just few instances of people anthropomorphizing the robot and
%interacting with it in a human-social way. More importantly, overall, these
%anthropomorphic interactions with the robot did not last but seemed to be
%somehow related to the novelty of the robot.

We categorized several interactions and actions with the robot as
anthropomorphic or human-social (in a broader sense). 
%Examples of what we categorized as somehow social and anthropomorphic
%(human-like) interactions with the robot include \textit{watching} the robot
%for fun, giving a \textit{demo} of the robot to guests / visitors, taking a
%\textit{photo/video} of the robot, \textit{personalizing} the robot, giving a
%\textit{name} to the robot, \textit{talking} to the robot using direct speech,
%\textit{playing} with the robot, and making \textit{adjustments} to the home
%due to the robot. 
Figure \ref{fig:roomba-activities} shows how many of the 9 households engaged
in these activities during the 6 months when the study was conducted.

\begin{figure}[hb]
    \centering
    \includegraphics[width=0.5\columnwidth]{roomba-activities.png}
    \caption{Activities with the robot over time, n = 9 households. This bar
        chart shows how many of the 9 households carried out different kinds of
        anthropomorphic / social activities with and related to the robot at 3
        time points during 6 months. Activities include \texttt{watching} the
        robot for fun, giving a \texttt{demo} of the robot to guests / visitors,
        \texttt{personalizing} the robot, giving a \texttt{name} to the robot,
        \texttt{talking} to the robot, \texttt{playing} with the robot.
        Participants were asked to report on these activities at 3 time points
        after the robot was deployed in their home: after 2 weeks, after 2 months,
        and after 6 months.}

@@

We also studied people's perception of the robot, and how it evolved over time
(see Figure \ref{fig:roomba-perception}). The main idea was to find out which of
the perceived attributes could be related to how far the robot was integrated in
the household as a cleaning tool. Concerning anthropomorphism, we included
\textit{perceived intelligence} and experienced \textit{fun} when using the
robot as indicators for the perception of human-likeness in the robot.\fixme{JF: Severin, you also mentioned impact as one factor. Please include it here below as another paragraph.}

\begin{figure}[b]
    \centering
    \includegraphics[width=0.5\columnwidth]{roomba-perception.pdf}

%    \begin{tikzpicture}[scale=0.5, y = 2cm]
%    %axis
%    \draw (0,0) -- coordinate (x axis mid) (20,0);
%    \draw (0,0) -- coordinate (y axis mid) (0,7);
%    %ticks
%    \foreach \x in {0,4,...,20}
%        \draw (\x,1pt) -- (\x,-3pt);
%    \foreach \y in {1,...,7} 
%        \draw[black!20] (20,\y) -- (-3pt,\y);
%    \foreach \y in {1,...,7} 
%        \draw (-3pt, \y) node[anchor=east] {\y}; 
%    %labels
%    \node[below=0.8cm] at (x axis mid) {People's perception ($\sigma=15$)};
%    \node at (2, -0.4) {\tiny expectations};
%    \node[align=center] at (6, -0.4) {\tiny after\\ \tiny first usage};
%    \node[align=center] at (10, -0.4) {\tiny after\\ \tiny 2 weeks};
%    \node[align=center] at (14, -0.4) {\tiny after\\ \tiny 2 months};
%    \node[align=center] at (18, -0.4) {\tiny after\\ \tiny 6 months};
%
%    \draw[dashed, black!50] (6,0) -- (6,7) node[sloped, below, pos=0.25] {\tiny novelty effect};
%
%    \draw[ultra thick, red!50] (2,5.6) -- (6,6.1) -- (10,6.3) -- (14,6.3) -- (18,6.1) node at (15,6.5) {\tiny \it ease of use};
%    \draw[ultra thick, purple] (2,4.9) -- (6,5.2) -- (10,5.1) -- (14,4.5) -- (18,4.2) node at (3,4.7) {\tiny \it usefulness};
%    \draw[ultra thick, blue!50] (2,4) -- (6,4.1) -- (10,4.6) -- (14,4.1) -- (18,4.3) node at (16,4) {\tiny \it fun};
%    \draw[ultra thick, green!50] (2,3.9) -- (6,4.1) -- (10,4.2) -- (14,3.8) -- (18,3.3) node at (19,3.5) {\tiny \it intelligence};
%    \draw[ultra thick, orange] (2,3.5) -- (6,4.2) -- (10,4) -- (14,3.2) -- (18,2.8) node at (14,2.7) {\tiny \it impact};
%
%    \draw[ultra thick, dashed] (2,5.2) -- (6,5.2) -- (10,5) -- (14,4.9) -- (18,4.5) node at (14,5.3) {\tiny \it overall};
%
%    \end{tikzpicture}

@@

@@

\begin{enumerate}
    	\item People engage in a more social and human-like way with a robotic pet compared to a functional robot and a tablet computer, therefore there will be a higher amount of anthropomorphic language in the AIBO and Roomba forum, compared to the iPad forum.

@@

\begin{figure}[b]
    \centering
    \includegraphics[width=0.5\columnwidth]{aibo-roomba-ipad_2.jpg}
    \caption{The Sony AIBO robotic dog (left), the iRobot Roomba vacuum-cleaning robot (middle), and the Apple Inc. iPad tablet computer (right) were compared in terms of how far users of these devices use anthropomorphic expressions in discussions about their technology / robot in online discussion forums.}
    \label{fig:aibo-roomba-ipad}
\end{figure}

@@

\begin{itemize}
    \item \textit{Life-likeness}, such as being alive or having (parts of) a
        body, \eg \textit{``she can be reborn since you copied her memory
        stick''};
    \item \textit{Emotional states, empathy} or the technology having a
        \textit{feeling}, \eg \textit{``oh no, poor AIBO''};
    \item \textit{Gender, personality} or the technology having an
        \textit{intention}, \eg Roomba \textit{``seems to like to hang around
        under the sofa''};
    \item When the author gave the technology a \textit{name}, \eg
        \textit{``whenever I show Java his pink ball''};
    \item Being \textit{socially integrated}, such as being a family member, \eg
        \textit{``I'm considering adding a Roomba to the family''};
    \item \textit{Metaphorical ways}, \eg Roomba \textit{``sings its victory
        song when it finishes and docks''}.
\end{itemize}

In the next paragraphs, we present findings in respect of our two hypotheses:
anthropomorphic language related to the device, and anthropomorphic language
related to the topic of the forum post.

%Overall, our analysis revealed that in all three forums, discussions about
%\textit{technical} aspects of the respective device were most frequent (67.7~\%
%of all 1208 categorized segments). However there were huge differences between
%the three devices: \textit{technical} aspects made up 79.9~\% of the discussion
%in the Roomba forum, compared to 69.9~\% in the AIBO forum, and 55.2~\% in the
%iPad forum. Another fairly frequently discussed topic concerned the
%\textit{usage} of the respective device, thus people describing how and for what
%they use their technology. This topic made up 21.8~\% of all posts, and was most
%frequent in the iPad forum (35.7~\%), followed by the Roomba forum (15.9~\%),
%and the AIBO forum (10.7~\%). Last but not least, we defined \textit{relation}
%as the third topic of interest. Discussions in which people described their
%relation to the respective device made up 10.5~\% of all posts, and was most
%frequent in the AIBO forum (19.4~\%), followed by the iPad forum (9.1~\%), and
%the Roomba forum (4.2~\%). \hl{(if it is getting too much we can leave out this
%paragraph, it's not crucial, and doesn't say anything about anthropomorphism
%directly)}

@@

13 groups of each two children (n = 26) who were all between 4-5 years old and
knew each other, were invited to play with the Ranger
robot toy box \citep{fink2014which} to assemble a domino together. The robot
was operated by one of the experimenters using the wizard-of-oz technique.
Ranger was supposed to transport domino tiles between the two children and it
would do its job correctly for the first 7 runs, bringing each one domino tile to the child. Then, however, the
robot started misbehaving by not coming over to the child who was calling the 
robot. Three different experimental conditions were explored using three different 
types of misbehavior from the robot side:
Either the robot was \textit{disobeying} and did not come over at all, or 
the robot got \textit{lost} on its way to the child, or the robot did a 
\textit{mistake} by first going in another direction but then turning and 
coming over to the child who had called it.
According to these three different types of unexpected robot behavior, we 
expected to observe variations of how far the children perceived the 
robot in terms of intentionality, and cognitive abilities.
